---
title: "Xuejian(Susan) Wang"
---


I’m an educator turning data scientist. For over a decade, I taught English and math in the U.S. and China, while also designing programs, managing teams, and improving learning outcomes through data.

My experiences in education inspired me to focus more deeply on data and problem-solving, which led me to pursue a Master’s in Statistics (Data Science Track) at Baruch College. Today, I’m building strong analytical and technical skills to connect my teaching background with meaningful, data-driven insights.

My mission: to use DATA to uncover patterns, solve problems, and create practical solutions that make a positive impact.

Welcome to connect me at my [Linkedin](https://www.linkedin.com/in/xuejian-susan-wang-987881225/) and [github](https://github.com/BH2xj).


```{r}
#| echo: false
#| message: false
#| warning: false

if(!require("leaflet")){
    options(repos=c(CRAN="https://cloud.r-project.org"))
    install.packages("leaflet")
    stopifnot(require("leaflet"))
}

baruch_longitude <- -73.98333
baruch_latitude  <- +40.74028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom=17) |>
  addPopups(baruch_longitude, baruch_latitude, 
            "I am a Master's student at <b>Baruch College</b>!")
```


```{r}
#| include: false
1+1
```
--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`